---
title: "Conceptual models for quantifying potential warming effects on ∆14C respired"
author: "J. Beem-Miller"
date: "2024-06-27"
output: html_document
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = c('cairo_pdf', 'png'))
options(scipen = 5)
# load read ess_dive fx
source("./utilities/read_ess-dive.fx.R")
```

```{r setup, include = FALSE}
library(ggplot2)
library(tidyr)
library(SoilR)
library(readxl)
library(ISRaD)
library(gt)
library(mpspline2)
library(dplyr)
library(forecast)
library(cowplot)
```

# Exp t0
- start dates of other candidate experiments listed here:
* B4W: 2006
* Blgt: 2013
* SPRUCE: 2015
* SWELTR: 2016

# Load atmospheric 14C record
- use Hua 2021 (data from SoilR)
```{r atm14c}
# use Hua2021 (SoilR package)
NHZone2 <- read_excel(
  "/Users/jeff/sra-frc/data/external/Hua_2021/S0033822221000953sup002.xls", sheet = 2, skip = 5,
  col_names = c("Year.AD", "mean.Delta14C", "sd.Delta14C", "mean.F14C", "sd.F14C")) %>%
  data.frame

# forecast data through 2025
yrs <- seq(2000, 2019.25, by = 1/4) # final sequence of years by quarters
nz2 <- spline(NHZone2[ , c(1, 4)], xout = yrs) # quarterly spline interpolation of fm data
nhz2 <- ts(nz2$y, start = 2000, freq = 4) # Transformation into a time-series object
m <- ets(nhz2) # Fits an exponential smoothing state space model to the time series
f2 <- forecast(m, h = 5.5 * 4) # Uses the fitted model to forecast  years into the future

# bind pre and post-bomb curves; add forecasted data
atm14c <- rbind(
  bind.C14curves(IntCal20, NHZone2, "AD"),
  data.frame(Year.AD = seq(tsp(f2$mean)[1], tsp(f2$mean)[2],
                           by = 1 / tsp(f2$mean)[3]),
             Delta14C = suppressWarnings(
               convert_fm_d14c(
                 fm = as.numeric(f2$mean), 
                 obs_date_y = seq(2019.75, 2022, by = .25), 
                 verbose = FALSE)),
             Sigma = NA)
)

# filter to 1900-2022 and calc annual averages
Datm <- data.frame(Date = seq(1900.5, 2024.5), d14c = NA)
for (i in seq_along(Datm$Date)) {
  ix <- which(atm14c$Year.AD >= Datm[i, "Date"] & atm14c$Year.AD < Datm[i, "Date"] + 1)
  Datm[i, "d14c"] <- mean(atm14c[ix, "Delta14C"], na.rm = TRUE)
}
```

# Utility fxs
```{r utilities}
# steady-state stock calc fx
calc.soc <- function(A, in_vector) {
  (-1 * solve(A) %*% in_vector)
}

# lambda (true half-life of 14C)
lambda <- 1 / 8267

# get initial fm value from k (pre-bomb)
fm <- function (k) {
  k / (k + lambda)
}

# mod fun 2pp
mod.fun.2pp <- function(
    t0 = 1900.5, 
    In = 1, 
    pars,
    scenario = "control",
    ctl.ls = NULL, ...) {
  
  # subset atm14c
  ATM14c <- Datm[Datm$Date > t0 - 1, ]
  
  # get input vector
  IN_VECTOR <- c(In * pars[3], In * (1 - pars[3]))
  
  # init. cond. 
  if (scenario == "control") {
    A <- -diag(pars[1:2])
    SOC <- calc.soc(A, IN_VECTOR)[,1]
    F0_Delta14C <- Delta14C_from_AbsoluteFractionModern(fm(pars[1:2]))
  } else {
    SOC <- ctl.ls[[2]][t0 - 1900.5, ]
    F0_Delta14C <- ctl.ls[[1]] %>%
      filter(years == t0) %>%
      filter(pool == "fast" | pool == "slow") %>%
      select(d14C) %>%
      unlist
  }
  
  mod <- TwopParallelModel14(
    t = ATM14c[,1],
    ks = pars[1:2],
    C0 = SOC,
    F0_Delta14C = F0_Delta14C,
    In = In,
    gam = pars[3],
    inputFc = ATM14c,
    lambda = lambda,
    lag = 0,
    pass = FALSE) 

  # get data
  C14m <- getF14C(mod)
  C14p <- getF14(mod) 
  C14r <- getF14R(mod)
  Ctot <- getC(mod)
  
  c14 <- data.frame(
    years = rep(ATM14c$Date, (ncol(C14p) + 3)),
    d14C = c(c(C14p),
             C14m,
             C14r,
             ATM14c$d14c),
    pool = rep(c("fast", "slow", "bulkC", "respiration", "atm"), 
               each = nrow(C14p)),
    scenario = scenario) %>% 
    distinct
  list(c14, Ctot)
}
```

# Control scenario
- two pool parallel model (2pp) selected for tractability
- first order kinetics
- no transfers between pools; "gamma" controls proportion of inputs entering fast pool
- "fast" pool tau = 5 y; "slow" pool tau = 100 y; gam = 0.9
- steady-state SOC stocks calculated using parameter set with input = 1 (unitless)
- time zero set to 1900
- time zero ∆14C calculated from k values

## ctl

```{r control}
## pars
In <- 1
k1.2pp <- .2
k2.2pp <- .01
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod w/ default input = 1
ctl.ls <- mod.fun.2pp(pars = pars)

# calc TT & SA
u <- c(In * pars[3], In * (1 - pars[3]))
A <- -diag(pars[1:2])
ctl.TT <- transitTime(A, u)
ctl.SA <- systemAge(A, u, a = seq(0, 500))
```

## 2xIn

```{r control-2xI}
## pars
In <- 2
k1.2pp <- .2
k2.2pp <- .01
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod w/ default input = 1
ctl.2x1.ls <- mod.fun.2pp(t0 = 1991.5, In = In, pars = pars, scenario = "ctl, 2xI", ctl.ls = ctl.ls)

# NB: TT & SA unchanged by inputs
```


# Warming scenarios
- all scenarios will consider a warming treatment of 4 °C
- Q10s will be calculated on the basis of this temperature difference, and decomposition rates will be adjusted by possible Q10 values
- initial values for ∆14C and SOC will be obtained from control model for desired experiment t0

## f=2; '91

```{r warm-q10-f2-s1}
## pars
In <- 1
k1.2pp <- .2 * 2^(4/10)
k2.2pp <- .01
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod
wrm.s10.f20.91.ls <- mod.fun.2pp(t0 = 1991.5, pars = pars, scenario = "wrm, s = 1.0, f = 2.0", ctl.ls = ctl.ls)

# calc ∆14C dif
ctl.ls[[1]][ctl.ls[[1]][["years"]] == 2024.5, "d14C"] - wrm.s10.f20.91.ls[[1]][wrm.s10.f20.91.ls[[1]][["years"]] == 2024.5, "d14C"]

# calc TT & SA
u <- c(In * pars[3], In * (1 - pars[3]))
A <- -diag(pars[1:2])
wrm.s10.f20.91.TT <- transitTime(A, u)
wrm.s10.f20.91.SA <- systemAge(A, u, a = seq(0, 500))
```

## f=2, s=2; '91

```{r warm-q10-f2-s2}
## pars
In <- 1
k1.2pp <- .2 * 2^(4/10)
k2.2pp <- .01 * 2^(4/10)
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod
wrm.s20.f20.91.ls <- mod.fun.2pp(t0 = 1991.5, pars = pars, scenario = "wrm, s = 2.0, f = 2.0", ctl.ls = ctl.ls)

# calc ∆14C dif
ctl.ls[[1]][ctl.ls[[1]][["years"]] == 2024.5, "d14C"] - wrm.s20.f20.91.ls[[1]][wrm.s20.f20.91.ls[[1]][["years"]] == 2024.5, "d14C"]

# calc TT & SA
u <- c(In * pars[3], In * (1 - pars[3]))
A <- -diag(pars[1:2])
wrm.s20.f20.91.TT <- transitTime(A, u)
wrm.s20.f20.91.SA <- systemAge(A, u, a = seq(0, 500))
```

## f=2, s=2;' 91; 2xI

```{r warm-q10-f2-s2-2xI}
## pars
In <- 2
k1.2pp <- .2 * 2^(4/10)
k2.2pp <- .01 * 2^(4/10)
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod
wrm.s20.f20.2xI.ls <- mod.fun.2pp(t0 = 1991.5, In = In, pars = pars, scenario = "wrm, s = 2.0, f = 2.0, 2xI", ctl.ls = ctl.ls)

# calc ∆14C dif
ctl.ls[[1]][ctl.ls[[1]][["years"]] == 2024.5, "d14C"] - wrm.s20.f20.2xI.ls[[1]][wrm.s20.f20.2xI.ls[[1]][["years"]] == 2024.5, "d14C"]
``` 

It appears that while transit time and age do not change with increased inputs, the ∆14C becomes more depleted. Odd.

## f=2, s=2; '16

```{r warm-q10-f2-s2-16}
## pars
In <- 1
k1.2pp <- .2 * 2^(4/10)
k2.2pp <- .01 * 2^(4/10)
gam <- .9
pars <- c(k1.2pp, k2.2pp, gam)

## run mod
wrm.s20.f20.16.ls <- mod.fun.2pp(t0 = 2016.5, pars = pars, scenario = "wrm, s = 2.0, f = 2.0", ctl.ls = ctl.ls)
```

# TS Plots

```{r plot-fx}
p.fx <- function(df.ls, xlim = NULL, ylim = NULL, scenarios, lg.pos = "right", tSz = 14) {
  
  if (is.data.frame(df.ls)) {
    DF <- df.ls
  } else {
    DF <- bind_rows(df.ls) 
  }
  
  DF <- DF %>%
    mutate(scenario = factor(scenario, levels = scenarios)) 
  
  p <- ggplot(DF, aes(years, d14C, color = pool, linetype = scenario)) +
    geom_line() +
    scale_color_manual(
      name = "", 
      values = c("atm" = "gray", 
                 "bulkC" = "black",
                 "respiration" = "#8549c3", #6c36a2, #8549c3
                 "fast" = "#d8006c",
                 "slow" = "#006cd8")) +
    scale_linetype(name = "", drop = FALSE) +
    guides(
      color = guide_legend(ncol = 1),
      linetype = guide_legend(ncol = 1)) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          text = element_text(size = tSz),
          legend.position = lg.pos)
  
  if (!is.null(xlim)) {
    p <- p +
      scale_x_continuous(limits = xlim)
  }
  
  if (!is.null(ylim)) {
    p <- p +
      scale_y_continuous(limits = ylim)
  }
  p
}
```

## Control

```{r plot-ctl}
ctl.p <- p.fx(
  ctl.ls[[1]], 
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "wrm, s = 1.0, f = 2.0",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "none")
ggsave("/Users/jeff/eco-warm/src/plots/1991/ctl.2pp.png", plot = ctl.p,
       width = 5.56, height = 4.18, units = "in")
lg.ctl <- p.fx(
  ctl.ls[[1]],
  xlim = c(1991, 2025), ylim = c(-20, 200),
  scenarios = c("control",
                "wrm, s = 1.0, f = 2.0",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "right")
ggsave("/Users/jeff/eco-warm/src/plots/1991/lg.ctl.png", plot = lg.ctl,
       width = 5.56, height = 4.18, units = "in")
```

## ctl; 2xI

```{r plot-ctl}
ctl.2xI.p <- p.fx(
  list(ctl.ls[[1]], ctl.2x1.ls[[1]], wrm.s20.f20.2xI.ls[[1]]), 
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "ctl, 2xI",
                "wrm, s = 2.0, f = 2.0, 2xI"),
  lg.pos = "none")
ggsave("/Users/jeff/eco-warm/src/plots/1991/ctl.2xI.2pp.png", plot = ctl.2xI.p,
       width = 5.56, height = 4.18, units = "in")

lg.2xI.p <- p.fx(
  list(ctl.ls[[1]], ctl.2x1.ls[[1]], wrm.s20.f20.2xI.ls[[1]]), 
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "ctl, 2xI",
                "wrm, s = 2.0, f = 2.0, 2xI"),
  lg.pos = "right")
ggsave("/Users/jeff/eco-warm/src/plots/1991/lg.2xI.2pp.png", plot = lg.2xI.p,
       width = 5.56, height = 4.18, units = "in")
```


## f=2; '91

```{r plot-f2-91}
wrm.s1.f2.p <- p.fx(
  list(ctl.ls[[1]], wrm.s10.f20.91.ls[[1]]),
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "wrm, s = 1.0, f = 2.0",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "none")
ggsave("/Users/jeff/eco-warm/src/plots/1991/wrm.s1.f2.2pp.png", plot = wrm.s1.f2.p, width = 5.56, height = 4.18, units = "in")
```

## f=2, s=2; '91

```{r plot-f2-s2-91}
wrm.s2.f2.p <- p.fx(
  list(ctl.ls[[1]], wrm.s10.f20.91.ls[[1]], wrm.s20.f20.91.ls[[1]]),
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "wrm, s = 1.0, f = 2.0",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "none")
ggsave("/Users/jeff/eco-warm/src/plots/1991/wrm.s2.f2.2pp.png", plot = wrm.s2.f2.p, width = 5.56, height = 4.18, units = "in")
```

## f=2, s=2; '16

```{r plot-f2-s2-16}
wrm.s2.f2.16.p <- p.fx(
  list(ctl.ls[[1]], wrm.s20.f20.16.ls[[1]]),
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "none")
ggsave("/Users/jeff/eco-warm/src/plots/2016/wrm.s2.f2.16.2pp.png", plot = wrm.s2.f2.16.p, width = 5.56, height = 4.18, units = "in")

lg.16.p <- p.fx(
  list(ctl.ls[[1]], wrm.s20.f20.16.ls[[1]]),
  xlim = c(1991, 2025), ylim = c(-20, 200), 
  scenarios = c("control",
                "wrm, s = 2.0, f = 2.0"),
  lg.pos = "right")
ggsave("/Users/jeff/eco-warm/src/plots/2016/lg.16.png", plot = lg.16.p, width = 5.56, height = 4.18, units = "in")
```

# SOC stocks
## Q10 = 2; 2xI; 1991

```{r plot-soc}
SOC.ls <- list(ctl.2x1.ls, wrm.s20.f20.91.ls, wrm.s20.f20.2xI.ls)
soc.dif.df <- bind_rows(lapply(seq_along(SOC.ls), function(i) {
  df <- data.frame(SOC.ls[[i]][[2]] - ctl.ls[[2]][1:nrow(SOC.ls[[i]][[2]]), ])
  df <- cbind(seq(0, nrow(df) - 1, 1), df)
  names(df) <- c("years", "f_dif", "s_dif")
  df$scenario <- SOC.ls[[i]][[1]][1:nrow(df), "scenario"]
  df %>%
    pivot_longer(cols = c("f_dif", "s_dif"), names_to = "pool", values_to = "SOC_dif")
}))

# plot
soc.p <- ggplot(soc.dif.df, aes(years, SOC_dif, color = pool, linetype = scenario)) +
  geom_line() +
  scale_color_manual(
    values = c("f_dif" = "#d8006c",
               "s_dif" = "#006cd8")) +
  theme_bw() +
  theme(panel.grid = element_blank())
ggsave("/Users/jeff/eco-warm/src/plots/1991/soc.png", plot = soc.p, height = 4.18, width = 7, unit = "in")
```

# Age & transit time

```{r TT-plots}
tt.ls <- list(ctl = ctl.TT, s10.f20 = wrm.s10.f20.91.TT, s20.f20 = wrm.s20.f20.91.TT)
tt.df <- bind_rows(lapply(seq_along(tt.ls), function(i) {
  data.frame(density = tt.ls[[i]][["transitTimeDensity"]], mod = names(tt.ls)[[i]]) %>%
    mutate(years = seq(0, 100, 1))
}))
tt.q <- bind_rows(lapply(tt.ls, function(x) {
  data.frame(years = x[["quantiles"]], quantile = c("5", "50", "95"))
  }), .id = "mod")

tt.p <- ggplot(tt.df, aes(years, density, linetype = mod)) +
  geom_line() +
  geom_vline(
    data = tt.q, 
    aes(xintercept = years, linetype = mod, color = quantile)) +
  scale_x_continuous(limits = c(0, 75)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 14))
ggsave("/Users/jeff/eco-warm/src/plots/tt.png", plot = tt.p, height = 4.18, width = 7, unit = "in")
```

```{r SA-plots}
sa.ls <- list(ctl = ctl.SA, s10.f20 = wrm.s10.f20.91.SA, s20.f20 = wrm.s20.f20.91.SA)
sa.df <- bind_rows(lapply(seq_along(sa.ls), function(i) {
  data.frame(density = sa.ls[[i]][["systemAgeDensity"]], mod = names(sa.ls)[[i]]) %>%
    mutate(years = seq(0, 500, 1))
}))
sa.q <- bind_rows(lapply(sa.ls, function(x) {
  data.frame(years = x[["quantilesSystemAge"]], quantile = c("5", "50", "95"))
  }), .id = "mod")

sa.p <- ggplot(sa.df, aes(years, density, linetype = mod)) +
  geom_line() +
  geom_vline(
    data = sa.q, 
    aes(xintercept = years, linetype = mod, color = quantile)) +
  scale_x_continuous(limits = c(0, 275)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 14))
ggsave("/Users/jeff/eco-warm/src/plots/sa.png", plot = sa.p, height = 4.18, width = 7, unit = "in")

# lapply(sa.ls, "[[", "meanSystemAge")
```